---
title: "Fantasy Football Make Models"
output: html_document
---


####################################################################################
#Chapter 1: Setting up the script
####################################################################################


Clear Everything from memory 
```{r, echo=FALSE}
rm(list=ls())
```

Set wd, install packages and disable scientific notation
```{r setup, include=FALSE, echo=FALSE}
# Set up working directory
path <- paste("/Users/louisgraham/Dropbox/Fantasy Football")
knitr::opts_knit$set(root.dir = normalizePath(path))

# Prevent scientific notation
knitr::opts_knit$set(options(scipen=999))

# Install and load all the packages that will be used for analysis
pkgs <- c("caret", 'caretEnsemble', "tidyverse", "lubridate", "data.table", "dplyr", "readxl", "ggplot2",
          "stringi", "adagio", "lpSolve")

miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] # vector of missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}
invisible(lapply(pkgs,library,character.only=TRUE))

# Clear our memory by removing objects that are no longer needed.
rm(miss_pkgs, pkgs, path)
```


Open data 
```{r, echo=FALSE, message=FALSE}
players_for_prediction <- readRDS("Data/Cleaned Data/Players historical data.rds")
```

####################################################################################
#Chapter 2: Clean into databases for prediction
####################################################################################

Add data for the quarter of the season
```{r, echo=FALSE}
players_for_prediction <- players_for_prediction %>%
  mutate(quarter = ifelse(round <= 10, "first",
                          ifelse(round <= 19, "second",
                                 ifelse(round <= 29, "third", "fourth"))))
```

Calculate own team, other team and relative difficulty
```{r, echo=FALSE}
players_for_prediction <- players_for_prediction %>%
  mutate(own_team_difficulty = ifelse(season %in% c("18-19", "19-20", "20-21"), ifelse(was_home == TRUE, 
                                                                                       elo_team_h, elo_team_a),
                                      elo_team_h),
         other_team_difficulty = ifelse(season %in% c("18-19", "19-20", "20-21"), ifelse(was_home == TRUE,
                                                                                         elo_team_a, elo_team_h),
                                        elo_team_a)) %>%
  mutate(relative_difficulty = other_team_difficulty/own_team_difficulty)
```

Clean home or away
```{r, echo=FALSE}
players_for_prediction <- players_for_prediction %>%
  mutate(home = ifelse(was_home == TRUE, 1, 0))
```

Keep only the variables needed
```{r, echo=FALSE}
players_for_prediction <- players_for_prediction %>%
  select(ict_index, minutes, round, quarter, selected, total_points, value, name, season,
         relative_difficulty, home, penalties_order, penalties_dummy, 
         corner_order, corner_dummy, position)
```

Change selected, ict and value into relative amount (e.g. 0 for least selected, 1 for most) for that round
```{r, echo=FALSE}
players_for_prediction <- players_for_prediction %>%
  group_by(season, round) %>%
  mutate(selected = selected/max(selected),
         ict_index = ict_index/max(ict_index),
         value = value/max(value)) %>% ungroup()
```

For minutes, total_points, ICT index and value, take lagged
```{r, echo=FALSE}
players_for_prediction <- players_for_prediction %>%
  #Make season a number
  mutate(season = as.numeric(substr(season, 1, 2))) %>%
  #Group by player, and arrange by season and round
  group_by(name) %>%
  arrange(season, round) %>%
  #Take lagged variables
  mutate(minutes_lag1 = lag(minutes,1), minutes_lag2 = lag(minutes,2),
         minutes_lag3 = lag(minutes,3), minutes_lag4 = lag(minutes,4),
         minutes_lag5 = lag(minutes,5), minutes_lag6 = lag(minutes,6),
         ict_index_lag1 = lag(ict_index,1), ict_index_lag2 = lag(ict_index,2),
         ict_index_lag3 = lag(ict_index,3), ict_index_lag4 = lag(ict_index,4),
         ict_index_lag5 = lag(ict_index,5), ict_index_lag6 = lag(ict_index,6),
         total_points_lag1 = lag(total_points,1), total_points_lag2 = lag(total_points,2),
         total_points_lag3 = lag(total_points,3), total_points_lag4 = lag(total_points,4),
         total_points_lag5 = lag(total_points,5), total_points_lag6 = lag(total_points,6),
         value_lag1 = lag(value,1), value_lag2 = lag(value,2),
         value_lag3 = lag(value,3), value_lag4 = lag(value,4),
         value_lag5 = lag(value,5), value_lag6 = lag(value,6)) %>%
  mutate(value_change1 = 100*(value - value_lag1)/value,
         value_change2 = 100*(value - value_lag2)/value,
         value_change3 = 100*(value - value_lag3)/value,
         value_change4 = 100*(value - value_lag4)/value,
         value_change5 = 100*(value - value_lag5)/value,
         value_change6 = 100*(value - value_lag6)/value) %>% ungroup()
```

Remove first few games in a season
```{r, echo=FALSE}
players_for_prediction <- players_for_prediction %>%
  filter(round > 6)
```

Prep for modelling
```{r, echo=FALSE}
myControl <- trainControl(
  method = "repeatedcv", number=5, repeats=1,
  classProbs = FALSE, 
  verboseIter = FALSE,
  savePredictions = TRUE, 
  trim = TRUE
)
seed_object <- 3734
```

####################################################################################
#Chapter 3: Run prediction model for 6 lag
####################################################################################

Create dataset with skipped rows, and create model
```{r, echo=FALSE}
#Model
formula_lag6 <- as.formula(total_points ~ selected + relative_difficulty + home + 
                        minutes_lag1 + minutes_lag2 + minutes_lag3 +
                        minutes_lag4 + minutes_lag5 + minutes_lag6 +
                        ict_index_lag1 + ict_index_lag2 + ict_index_lag3 +
                        ict_index_lag4 + ict_index_lag5 + ict_index_lag6 +
                        total_points_lag1 + total_points_lag2 + total_points_lag3 +
                        total_points_lag4 + total_points_lag5 + total_points_lag6 +
                        value + value_change6 + penalties_dummy +
                        penalties_order*penalties_dummy + corner_dummy + corner_order*corner_dummy +
                        position + quarter)
#Model
tune_grid <- expand.grid(alpha = seq(0.99, 1, by = 0.005), lambda = seq(0, 0.1, by = 0.0025))
set.seed(seed_object)
model_glmnet_lag6 <- train(
  formula_lag6, data = players_for_prediction,
  metric = "RMSE",
  method = "glmnet",
  tuneGrid = tune_grid,
  trControl = myControl,
  na.action = na.omit
)
print(model_glmnet_lag6$bestTune)
rm(formula_lag6, tune_grid)
```

####################################################################################
#Chapter 4: Run prediction model for 5 lag
####################################################################################

Create dataset with skipped rows, and create model
```{r, echo=FALSE}
#Model
formula_lag5 <- as.formula(total_points ~ selected + relative_difficulty + home + 
                        minutes_lag1 + minutes_lag2 + minutes_lag3 +
                        minutes_lag4 + minutes_lag5 +
                        ict_index_lag1 + ict_index_lag2 + ict_index_lag3 +
                        ict_index_lag4 + ict_index_lag5 +
                        total_points_lag1 + total_points_lag2 + total_points_lag3 +
                        total_points_lag4 + total_points_lag5 +
                        value + value_change5 + penalties_dummy +
                        penalties_order*penalties_dummy + corner_dummy + corner_order*corner_dummy +
                        position + quarter)
#Model
tune_grid <- expand.grid(alpha = seq(0.99, 1, by = 0.005), lambda = seq(0, 0.1, by = 0.0025))
set.seed(seed_object)
model_glmnet_lag5 <- train(
  formula_lag5, data = players_for_prediction,
  metric = "RMSE",
  method = "glmnet",
  tuneGrid = tune_grid,
  trControl = myControl,
  na.action = na.omit
)
print(model_glmnet_lag5$bestTune)
rm(formula_lag5, tune_grid)
```

####################################################################################
#Chapter 5: Run prediction model for 4 lag
####################################################################################

Create dataset with skipped rows, and create model
```{r, echo=FALSE}
#Model
formula_lag4 <- as.formula(total_points ~ selected + relative_difficulty + home + 
                        minutes_lag1 + minutes_lag2 + minutes_lag3 +
                        minutes_lag4 +
                        ict_index_lag1 + ict_index_lag2 + ict_index_lag3 +
                        ict_index_lag4 +
                        total_points_lag1 + total_points_lag2 + total_points_lag3 +
                        total_points_lag4 +
                        value + value_change4 + penalties_dummy +
                        penalties_order*penalties_dummy + corner_dummy + corner_order*corner_dummy  +
                        position + quarter)
#Model
tune_grid <- expand.grid(alpha = seq(0.99, 1, by = 0.005), lambda = seq(0, 0.1, by = 0.0025))
set.seed(seed_object)
model_glmnet_lag4 <- train(
  formula_lag4, data = players_for_prediction,
  metric = "RMSE",
  method = "glmnet",
  tuneGrid = tune_grid,
  trControl = myControl,
  na.action = na.omit
)
print(model_glmnet_lag4$bestTune)
rm(formula_lag4, tune_grid)
```

####################################################################################
#Chapter 6: Run prediction model for 3 lag
####################################################################################

Create dataset with skipped rows, and create model
```{r, echo=FALSE}
#Model
formula_lag3 <- as.formula(total_points ~ selected + relative_difficulty + home + 
                        minutes_lag1 + minutes_lag2 + minutes_lag3 +
                        ict_index_lag1 + ict_index_lag2 + ict_index_lag3 +
                        total_points_lag1 + total_points_lag2 + total_points_lag3 +
                        value + value_change3 + penalties_dummy +
                        penalties_order*penalties_dummy + corner_dummy + corner_order*corner_dummy  +
                        position + quarter)
#Model
tune_grid <- expand.grid(alpha = seq(0.99, 1, by = 0.005), lambda = seq(0, 0.1, by = 0.0025))
set.seed(seed_object)
model_glmnet_lag3 <- train(
  formula_lag3, data = players_for_prediction,
  metric = "RMSE",
  method = "glmnet",
  tuneGrid = tune_grid,
  trControl = myControl,
  na.action = na.omit
)
print(model_glmnet_lag3$bestTune)
rm(formula_lag3, tune_grid)
```

####################################################################################
#Chapter 7: Run prediction model for 2 lag
####################################################################################

Create dataset with skipped rows, and create model
```{r, echo=FALSE}
#Model
formula_lag2 <- as.formula(total_points ~ selected + relative_difficulty + home + 
                        minutes_lag1 + minutes_lag2 + 
                        ict_index_lag1 + ict_index_lag2 + 
                        total_points_lag1 + total_points_lag2 + 
                        value + value_change2 + penalties_dummy +
                        penalties_order*penalties_dummy + corner_dummy + corner_order*corner_dummy  +
                        position + quarter)
#Model
tune_grid <- expand.grid(alpha = seq(0.99, 1, by = 0.005), lambda = seq(0, 0.1, by = 0.0025))
set.seed(seed_object)
model_glmnet_lag2 <- train(
  formula_lag2, data = players_for_prediction,
  metric = "RMSE",
  method = "glmnet",
  tuneGrid = tune_grid,
  trControl = myControl,
  na.action = na.omit
)
print(model_glmnet_lag2$bestTune)
rm(formula_lag2, tune_grid)
```

####################################################################################
#Chapter 8: Run prediction model for 1 lag
####################################################################################

Create dataset with skipped rows, and create model
```{r, echo=FALSE}
#Model
formula_lag1 <- as.formula(total_points ~ selected + relative_difficulty + home + 
                        minutes_lag1 + 
                        ict_index_lag1 + 
                        total_points_lag1 + 
                        value + value_change1 + penalties_dummy +
                        penalties_order*penalties_dummy + corner_dummy + corner_order*corner_dummy  +
                        position + quarter)
#Model
tune_grid <- expand.grid(alpha = seq(0.99, 1, by = 0.005), lambda = seq(0, 0.1, by = 0.0025))
set.seed(seed_object)
model_glmnet_lag1 <- train(
  formula_lag1, data = players_for_prediction,
  metric = "RMSE",
  method = "glmnet",
  tuneGrid = tune_grid,
  trControl = myControl,
  na.action = na.omit
)
print(model_glmnet_lag1$bestTune)
rm(formula_lag1, tune_grid)
```

####################################################################################
#Chapter 9: Run prediction model for those with no game history
####################################################################################

Regression with ridge and lasso
```{r, echo=FALSE}
#Model
formula_nolag <- as.formula(total_points ~ selected + relative_difficulty + home +
         value + penalties_dummy + penalties_dummy*penalties_order + corner_dummy +
           corner_dummy*corner_order + position + quarter)
set.seed(seed_object)
tune_grid <- expand.grid(alpha = seq(0.98, 1, by = 0.005), lambda = seq(0, 0.1, by = 0.0025))
#Run
model_glmnet_no_lag <- train(
  formula_nolag, data = players_for_prediction,
  metric = "RMSE",
  method = "glmnet",
  trControl = myControl,
  tuneGrid = tune_grid,
  na.action = na.omit
)
print(model_glmnet_no_lag$bestTune) 
rm(formula_nolag, tune_grid)
```

####################################################################################
#Chapter 10: Read and compare models
####################################################################################

Compare models
```{r, echo=FALSE}
model_list <- list(lag6 = model_glmnet_lag6, lag5 = model_glmnet_lag5, lag4 = model_glmnet_lag4,
                   lag3 = model_glmnet_lag3, lag2 = model_glmnet_lag2, lag1 = model_glmnet_lag1,
                   no_lag = model_glmnet_no_lag)
resamples <- resamples(model_list)
summary(resamples)
bwplot(resamples, metric = "RMSE")
```

Explore variable importance
```{r, echo=FALSE}
varImp(model_glmnet_lag6)
varImp(model_glmnet_lag5)
varImp(model_glmnet_lag4)
varImp(model_glmnet_lag3)
varImp(model_glmnet_lag2)
varImp(model_glmnet_lag1)
varImp(model_glmnet_no_lag)
```

Explore coefficients
```{r, echo=FALSE}
coefficients <- as.matrix(coef(model_glmnet_lag4$finalModel, model_glmnet_lag4$bestTune$lambda))
```

Remove objects
```{r, echo=FALSE}
rm(model_list, resamples, myControl, coefficients)
```

Save models (in a memory light way by stripping out the heavy model objects that aren't needed)
```{r, echo=FALSE}
model_glmnet_lag6[c("trainingData", "pred", "control", "na.action")] <- NULL
model_glmnet_lag5[c("trainingData", "pred", "control", "na.action")] <- NULL
model_glmnet_lag4[c("trainingData", "pred", "control", "na.action")] <- NULL
model_glmnet_lag3[c("trainingData", "pred", "control", "na.action")] <- NULL
model_glmnet_lag2[c("trainingData", "pred", "control", "na.action")] <- NULL
model_glmnet_lag1[c("trainingData", "pred", "control", "na.action")] <- NULL
model_glmnet_no_lag[c("trainingData", "pred", "control", "na.action")] <- NULL
saveRDS(model_glmnet_lag6, "Models/Lasso_Ridge_Model_lag6.Rds")
saveRDS(model_glmnet_lag5, "Models/Lasso_Ridge_Model_lag5.Rds")
saveRDS(model_glmnet_lag4, "Models/Lasso_Ridge_Model_lag4.Rds")
saveRDS(model_glmnet_lag3, "Models/Lasso_Ridge_Model_lag3.Rds")
saveRDS(model_glmnet_lag2, "Models/Lasso_Ridge_Model_lag2.Rds")
saveRDS(model_glmnet_lag1, "Models/Lasso_Ridge_Model_lag1.Rds")
saveRDS(model_glmnet_no_lag, "Models/Lasso_Ridge_Model_No_Lag.Rds")
```

####################################################################################
#Chapter 11: Explore predictions
####################################################################################

Explore predictions
```{r, echo=FALSE}
players_for_prediction <-  players_for_prediction %>%
mutate(points_predicted_6lag = predict(model_glmnet_lag6, newdata = players_for_prediction, na.action = NULL),
       points_predicted_5lag = predict(model_glmnet_lag5, newdata = players_for_prediction, na.action = NULL),
       points_predicted_4lag = predict(model_glmnet_lag4, newdata = players_for_prediction, na.action = NULL),
       points_predicted_3lag = predict(model_glmnet_lag3, newdata = players_for_prediction, na.action = NULL),
       points_predicted_2lag = predict(model_glmnet_lag2, newdata = players_for_prediction, na.action = NULL),
       points_predicted_1lag = predict(model_glmnet_lag1, newdata = players_for_prediction, na.action = NULL),
       points_predicted_no_lag = predict(model_glmnet_no_lag, newdata = players_for_prediction, na.action = NULL))
```


Explore correlations between predictions and actual
```{r, echo=FALSE}
#Correlations
cor(players_for_prediction$points_predicted_6lag, players_for_prediction$total_points, use="complete.obs")
cor(players_for_prediction$points_predicted_5lag, players_for_prediction$total_points, use="complete.obs")
cor(players_for_prediction$points_predicted_4lag, players_for_prediction$total_points, use="complete.obs")
cor(players_for_prediction$points_predicted_3lag, players_for_prediction$total_points, use="complete.obs")
cor(players_for_prediction$points_predicted_2lag, players_for_prediction$total_points, use="complete.obs")
cor(players_for_prediction$points_predicted_1lag, players_for_prediction$total_points, use="complete.obs")
cor(players_for_prediction$points_predicted_no_lag, players_for_prediction$total_points, use="complete.obs")
#Graphs
plot(players_for_prediction$points_predicted_4lag, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$points_predicted_4lag))
plot(players_for_prediction$points_predicted_2lag, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$points_predicted_2lag))
plot(players_for_prediction$points_predicted_1lag, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$points_predicted_1lag))
plot(players_for_prediction$points_predicted_no_lag, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$points_predicted_no_lag))
```

Explore correlations between predictions, and variables
```{r, echo=FALSE}
#Graphs
plot(players_for_prediction$value, players_for_prediction$points_predicted_4lag)
abline(lm(players_for_prediction$points_predicted_4lag ~ players_for_prediction$value))
plot(players_for_prediction$selected, players_for_prediction$points_predicted_4lag)
abline(lm(players_for_prediction$points_predicted_4lag ~ players_for_prediction$selected))
plot(players_for_prediction$relative_difficulty, players_for_prediction$points_predicted_4lag)
abline(lm(players_for_prediction$points_predicted_4lag ~ players_for_prediction$relative_difficulty))
plot(players_for_prediction$ict_index_lag1, players_for_prediction$points_predicted_4lag)
abline(lm(players_for_prediction$points_predicted_4lag ~ players_for_prediction$ict_index_lag1))
plot(players_for_prediction$total_points_lag1, players_for_prediction$points_predicted_4lag)
abline(lm(players_for_prediction$points_predicted_4lag ~ players_for_prediction$total_points_lag1))
plot(players_for_prediction$minutes_lag1, players_for_prediction$points_predicted_4lag)
abline(lm(players_for_prediction$points_predicted_4lag ~ players_for_prediction$minutes_lag1))
plot(players_for_prediction$penalties_dummy, players_for_prediction$points_predicted_4lag)
abline(lm(players_for_prediction$points_predicted_4lag ~ players_for_prediction$penalties_dummy))
```


Explore correlations between actual, and variables
```{r, echo=FALSE}
#Graphs
plot(players_for_prediction$value, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$value))
plot(players_for_prediction$selected, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$selected))
plot(players_for_prediction$relative_difficulty, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$relative_difficulty))
plot(players_for_prediction$ict_index_lag1, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$ict_index_lag1))
plot(players_for_prediction$total_points_lag1, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$total_points_lag1))
plot(players_for_prediction$minutes_lag1, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$minutes_lag1))
plot(players_for_prediction$penalties_dummy, players_for_prediction$total_points)
abline(lm(players_for_prediction$total_points ~ players_for_prediction$penalties_dummy))
```






