---
title: "Fantasy Football Choose team"
output: html_document
---


STRATEGY
Part 1: Create models at player level using data from the last few seasons
1. Predict number of points in a game
2. Use threat of team, threat of other team (or relative threat of team), home or away, metrics for form (potentially including rising form), metrics for likelihood to play (e.g. time in last game)
3. Build a smaller model which only looks at team threat, ICT and cost for new people
Part 2: Use model to create team
1. Predict points for each of the first 5 games
2. Choose an 11 which maximises expected points for 90k, bearing in mind potential positions. Choose the remaining 5 similarly




####################################################################################
#Chapter 1: Setting up the script
####################################################################################


Clear Everything from memory 
```{r, echo=FALSE}
rm(list=ls())
```

Set wd, install packages and disable scientific notation
```{r setup, include=FALSE, echo=FALSE}
# Set up working directory
path <- paste("/Users/louisgraham/Dropbox/Fantasy Football")
knitr::opts_knit$set(root.dir = normalizePath(path))

# Prevent scientific notation
knitr::opts_knit$set(options(scipen=999))

# Install and load all the packages that will be used for analysis
pkgs <- c("caret", 'caretEnsemble', "tidyverse", "lubridate", "data.table", "dplyr", "readxl", "ggplot2",
          "stringi", "adagio", "lpSolve", 'httr', 'jsonlite')

miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] # vector of missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}
invisible(lapply(pkgs,library,character.only=TRUE))

# Clear our memory by removing objects that are no longer needed.
rm(miss_pkgs, pkgs, path)
```

####################################################################################
#Chapter 2: Prepare the dataframe looking 5 games ahead
####################################################################################


Open current players, fixtures, and historical players data
```{r, echo=FALSE, message=FALSE}
players_current <- readRDS("Cleaned Data/Current players.rds")
fixtures <- readRDS("Cleaned Data/All fixtures.rds")
players_historical <- readRDS("Cleaned Data/Players historical data.rds")
```

Keep only the variables needed for the modelling
```{r, echo=FALSE}
players_historical <- players_historical %>%
  select(minutes, round, total_points, value, name, season, ict_index, position,
         penalties_order, penalties_dummy, corner_order, corner_dummy)
```

For minutes, ict_index, total_points and value, take lagged 
```{r, echo=FALSE}
players_historical <- players_historical %>%
  #Make season a number
  mutate(season = as.numeric(substr(season, 1, 2))) %>%
  #Group by player, and arrange by season and round
  group_by(name) %>%
  arrange(season, round) %>%
  #Take lagged variables (bear in mind that the current is actually the first lag for the next game)
  mutate(minutes_lag1 = minutes, minutes_lag2 = lag(minutes,1),
         minutes_lag3 = lag(minutes,2), minutes_lag4 = lag(minutes,3),
         ict_index_lag1 = ict_index, ict_index_lag2 = lag(ict_index,1),
         ict_index_lag3 = lag(ict_index,2), ict_index_lag4 = lag(ict_index,3),
         total_points_lag1 = total_points, total_points_lag2 = lag(total_points,1),
         total_points_lag3 = lag(total_points,2), total_points_lag4 = lag(total_points,3),
         value_lag1 = value, value_lag2 = lag(value,1),
         value_lag3 = lag(value,2), value_lag4 = lag(value,3)) %>% ungroup() %>%
  #Drop the variables used for construction
  select(-minutes, -total_points, -value, -ict_index)
```

Clean and keep only the last row
```{r, echo=FALSE}
players_historical <- players_historical %>%
  group_by(name) %>%
  arrange(season, round) %>%
  #Keep the last row
  filter(row_number() == max(row_number())) %>% ungroup() %>%
  #Drop date variables
  select(-season, -round)
```

Merge players and historical players
```{r, echo=FALSE}
players <- left_join(players_current, players_historical, by = "name") 
rm(players_current, players_historical)
```

Merge with the fixtures data
```{r, echo=FALSE, message=FALSE}
players_away <- left_join(players, fixtures, by = c("team" = "team_a"))
players_away <- players_away %>%
  rename(own_team_difficulty = elo_team_a, 
         other_team_difficulty = elo_team_h) %>%
  mutate(home = 0) %>%
  select(-team_h)
players_home <- left_join(players, fixtures, by = c("team" = "team_h"))
players_home <- players_home %>%
  rename(own_team_difficulty = elo_team_h, 
         other_team_difficulty = elo_team_a) %>%
  mutate(home = 1) %>%
  select(-team_a)
players <- rbind(players_home, players_away) 
rm(players_home, players_away, fixtures)
```

Add data for the quarter of the season
```{r, echo=FALSE}
players <- players %>%
  mutate(quarter = ifelse(event <= 10, "first",
                          ifelse(event <= 19, "second",
                                 ifelse(event <= 29, "third", "fourth"))))
```

Calculate relative difficulty
```{r, echo=FALSE}
players$relative_difficulty <-  players$other_team_difficulty/players$own_team_difficulty
```

Adjust value, selected and ict variables
```{r, echo=FALSE, message=FALSE}
players <- players %>%
  mutate(value_change4 = 100*(now_cost - value_lag4)/now_cost,
         value_change3 = 100*(now_cost - value_lag3)/now_cost,
         value_change2 = 100*(now_cost - value_lag2)/now_cost,
         value_change1 = 100*(now_cost - value_lag1)/now_cost,
         selected = selected_by_percent/max(selected_by_percent),
         ict_index_lag1 = ict_index_lag1/max(ict_index_lag1, na.rm=T),
         ict_index_lag2 = ict_index_lag2/max(ict_index_lag2, na.rm=T),
         ict_index_lag3 = ict_index_lag3/max(ict_index_lag3, na.rm=T),
         ict_index_lag4 = ict_index_lag4/max(ict_index_lag4, na.rm=T),
         value = now_cost/max(now_cost)) %>%
  select(c(-selected_by_percent, -value_lag4, -value_lag3, -value_lag2, -value_lag1, -ict_index))
```

Shift name first
```{r, echo=FALSE, message=FALSE}
players <- players %>%
  select(name, everything())
```


####################################################################################
#Chapter 3: Make predictions
####################################################################################

Load in the models
```{r, echo=FALSE}
model_4_lag <- readRDS("Models/Lasso_Ridge_Model_lag4.Rds")
model_3_lag <- readRDS("Models/Lasso_Ridge_Model_lag3.Rds")
model_2_lag <- readRDS("Models/Lasso_Ridge_Model_lag2.Rds")
model_1_lag <- readRDS("Models/Lasso_Ridge_Model_lag1.Rds")
model_no_lag <- readRDS("Models/Lasso_Ridge_Model_No_Lag.Rds")
```

Run all models for as many people as possible, and choose 4lag, then 1lag, then nolag
```{r, echo=FALSE}
#Run 4lag
players$points_predicted_4_lag <- predict(model_4_lag, newdata = players, na.action = NULL)
#Run 3lag
players$points_predicted_3_lag <- predict(model_3_lag, newdata = players, na.action = NULL)
#Run 2lag
players$points_predicted_2_lag <- predict(model_2_lag, newdata = players, na.action = NULL)
#Run 1lag
players$points_predicted_1_lag <- predict(model_1_lag, newdata = players, na.action = NULL)   
#Run nolag
players$points_predicted_no_lag <- predict(model_no_lag, newdata = players, na.action = NULL)   
#Create points
players$points_predicted <- 
  ifelse(!is.na(players$points_predicted_4_lag), players$points_predicted_4_lag,
         ifelse(!is.na(players$points_predicted_3_lag), players$points_predicted_3_lag,
                ifelse(!is.na(players$points_predicted_2_lag), players$points_predicted_2_lag,
                       ifelse(!is.na(players$points_predicted_1_lag), players$points_predicted_1_lag,
                              players$points_predicted_no_lag))))
#Remove intermediate vars and models
players <- players %>%
  select(-points_predicted_4_lag, -points_predicted_3_lag, -points_predicted_2_lag, 
         -points_predicted_1_lag, -points_predicted_no_lag)
rm(model_4_lag, model_3_lag, model_2_lag, model_1_lag, model_no_lag)
```

Keep only the first five games
```{r, echo=FALSE}
players <- players %>%
  group_by(name) %>%
  arrange(kickoff_time) %>%
  filter(kickoff_time > as.Date(Sys.Date())) %>%
  filter(row_number() <= 5) %>% ungroup()
```

Take a weighted average of the first 5 games
```{r, echo=FALSE}
players <- players %>%
  group_by(name) %>%
  arrange(kickoff_time) %>%
  #Add in weight depending on how far away the game is (declining with time)
  mutate(weight = ifelse(row_number() == 1, 0.35, 
                         ifelse(row_number() == 2, 0.25,
                               ifelse(row_number() == 3, 0.2,
                                     ifelse(row_number() == 4, 0.12, 0.08))))) %>%
  #Create a weighted final points
  mutate(weighted_points = points_predicted * weight) %>%
  mutate(final_points = sum(weighted_points)) %>%
  #Keep one row only (this now gives a summary of expected points for the next games)
  filter(row_number() == 1) %>% ungroup() %>%
  #Rename the points predicted variable, as this is for the next game only. Note that several other variables are also for the next game. Keep these for visibility
  rename(points_predicted_next_game = points_predicted) %>%
  select(-weight, -weighted_points)
```


####################################################################################
#Chapter 5: Check current team
####################################################################################

Previous team (UPDATE THIS)
```{r, echo=FALSE}
previous_team <- data.frame(name = c("Emiliano Martinez", 
                                     "Kurt Zouma", "Ben Mee", "Benjamin Chilwell",
                                     "Jack Grealish",  "Bruno Miguel Borges Fernandes", "James Ward-Prowse", 
                                     "Heung-Min Son", 
                                     "Dominic Calvert-Lewin", "Harry Kane", "Patrick Bamford",
                                     "Alex McCarthy", "James Justin", "Vladimir Coufal", "Anwar El Ghazi"), 
                            first_11 = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0))
```

Merge team with the players data
```{r, echo=FALSE}
previous_team <- left_join(previous_team, players, by = "name")
```

Calculate best current 11 (at least 1 GK, 3 defenders and 1 forward)
```{r, echo=FALSE}
#Create objective function
objective.in <- previous_team$final_points
#Create matrix for requirements
const <- t(as.matrix(tibble(goal_keeper = ifelse(previous_team$position == "GK", 1, 0),
                            defender = ifelse(previous_team$position == "DEF", 1, 0),
                            midfield = ifelse(previous_team$position == "MID", 1, 0),
                            striker = ifelse(previous_team$position == "FWD", 1, 0),
                            ones = rep(1, nrow(previous_team)))))
#Create right hand side and inequality
rhs <- c(1, 3, 0, 1, 11)
direction  <- c("==", ">=", ">=", ">=", "==")
#Run, and make the team
optimum <-  lp(direction="max",  objective.in, const, direction,  rhs, all.bin = TRUE)
previous_team$first_11_new <- ifelse(optimum$solution==1, 1, 0)
rm(objective.in, const, rhs, direction, optimum)
```

Check expected statistics for current team
```{r, echo=FALSE}
sum(previous_team[previous_team$first_11_new == 1,]$now_cost)
sum(previous_team$now_cost)
sum(previous_team[previous_team$first_11_new == 1,]$final_points) + 
  max(previous_team[previous_team$first_11_new == 1,]$final_points)
```

####################################################################################
#Chapter 6: Select optimal team changes
####################################################################################

Amount left over (UPDATE THIS)
```{r, echo=FALSE}
amount_left_over <- 5
```

Rank potential switches. Loop through each player, and see the best person they could be transferred for given the cost and the requirement for team shape
```{r, echo=FALSE}
#Calculate number in each position. Don't go below 2 goalkeepers, 4 defenders and 2 strikers
number_by_position <- previous_team %>%
  group_by(position) %>%
  summarise(number = n())
#Loop through each player in the squad
switch <- data.frame(matrix(ncol = 21, nrow = 0))
for(i in 1:nrow(previous_team)){
  #Give details on the current player
  switch[i,1] <- previous_team$name[i]
  switch[i,2] <- previous_team$final_points[i]
  switch[i,3] <- previous_team$position[i]
  switch[i,4] <- previous_team$first_11_new[i]
  switch[i,5] <- previous_team$now_cost[i]
  #If a goalkeeper and there are 2 or fewer goalkeepers, has to be another goalkeeper
  if(switch[i,3] == "GK"  & number_by_position[number_by_position$position == "GK",]$number <= 2){
    df <- players[players$now_cost <= switch[i,5] + amount_left_over & players$position == "GK",]
  }
  #If a defender and there are 4 or fewer defenders, has to be another defender
  else if(switch[i,3] == "DEF"){#}  & number_by_position[number_by_position$position == "DEF",]$number <= 4){
    df <- players[players$now_cost <= switch[i,5] + amount_left_over & players$position == "DEF",]
  }
  #If a striker and there are 2 or fewer strikers, has to be another striker
  else if(switch[i,3] == "FWD"  & number_by_position[number_by_position$position == "FWD",]$number <= 2){
    df <- players[players$now_cost <= switch[i,5] + amount_left_over & players$position == "FWD",]
  }
  #Otherwise anyone
  else{
    df <- players[players$now_cost <= switch[i,5] + amount_left_over,]
  }
  #Get details on the replacement player
  switch[i,6] = df[df$final_points == sort(df$final_points, TRUE)[1] & !is.na(df$final_points),]$name
  switch[i,7] = df[df$final_points == sort(df$final_points, TRUE)[1]  & !is.na(df$final_points),]$final_points
  switch[i,8] = df[df$final_points == sort(df$final_points, TRUE)[1]  & !is.na(df$final_points),]$position
  switch[i,9] = df[df$final_points == sort(df$final_points, TRUE)[1]  & !is.na(df$final_points),]$now_cost
  switch[i,10] = df[df$final_points == sort(df$final_points, TRUE)[2] & !is.na(df$final_points),]$name
  switch[i,11] = df[df$final_points == sort(df$final_points, TRUE)[2]  & !is.na(df$final_points),]$final_points
  switch[i,12] = df[df$final_points == sort(df$final_points, TRUE)[2]  & !is.na(df$final_points),]$position
  switch[i,13] = df[df$final_points == sort(df$final_points, TRUE)[2]  & !is.na(df$final_points),]$now_cost
  switch[i,14] = df[df$final_points == sort(df$final_points, TRUE)[3] & !is.na(df$final_points),]$name
  switch[i,15] = df[df$final_points == sort(df$final_points, TRUE)[3]  & !is.na(df$final_points),]$final_points
  switch[i,16] = df[df$final_points == sort(df$final_points, TRUE)[3]  & !is.na(df$final_points),]$position
  switch[i,17] = df[df$final_points == sort(df$final_points, TRUE)[3]  & !is.na(df$final_points),]$now_cost
  switch[i,18] = df[df$final_points == sort(df$final_points, TRUE)[4] & !is.na(df$final_points),]$name
  switch[i,19] = df[df$final_points == sort(df$final_points, TRUE)[4]  & !is.na(df$final_points),]$final_points
  switch[i,20] = df[df$final_points == sort(df$final_points, TRUE)[4]  & !is.na(df$final_points),]$position
  switch[i,21] = df[df$final_points == sort(df$final_points, TRUE)[4]  & !is.na(df$final_points),]$now_cost
}
#Name the variables
x <- c("current_name", "current_points", "current_position", "current_first_11", "current_now_cost",
       "trade_name", "trade_points", "trade_position", "current_now_cost",
       "second_trade_name", "second_trade_points", "second_trade_position", "second_current_now_cost",
       "third_trade_name", "third_trade_points", "third_trade_position", "third_current_now_cost",
       "fourth_trade_name", "fourth_trade_points", "fourth_trade_position", "fourth_current_now_cost")
colnames(switch) <- x
rm(x, number_by_position, i, df)
#Calculate point benefit
switch$point_increase <- switch$trade_points - switch$current_points
switch$point_increase2 <- switch$second_trade_points - switch$current_points
switch$point_increase3 <- switch$third_trade_points - switch$current_points
switch$point_increase4 <- switch$fourth_trade_points - switch$current_points
#Choose person to transfer. Only include first team people
switch2 <- switch[switch$current_first_11 == 1,]
transfer_out <- switch[switch2$point_increase == max(switch2$point_increase),]$current_name
transfer_in <- switch[switch2$point_increase == max(switch2$point_increase),]$trade_name
rm(switch2)
```

Create new squad with the best change made
```{r, echo=FALSE}
new_team <- data.frame(name = previous_team$name) %>%
  filter(!name %in% transfer_out)
new_team <- rbind(new_team, data.frame(name = transfer_in))
new_team <- left_join(new_team, players, by = "name")
```

Create new team (at least 1 GK, 3 defenders and 1 forward)
```{r, echo=FALSE}
#Create objective function
objective.in <- new_team$final_points
#Create matrix for requirements
const <- t(as.matrix(tibble(goal_keeper = ifelse(new_team$position == "GK", 1, 0),
                            defender = ifelse(new_team$position == "DEF", 1, 0),
                            midfield = ifelse(new_team$position == "MID", 1, 0),
                            striker = ifelse(new_team$position == "FWD", 1, 0),
                            ones = rep(1, nrow(new_team)))))
#Create right hand side and inequality
rhs <- c(1, 3, 0, 1, 11)
direction  <- c("==", ">=", ">=", ">=", "==")
#Run, and make the team
optimum <-  lp(direction="max",  objective.in, const, direction,  rhs, all.bin = TRUE)
new_team$first_11_new <- ifelse(optimum$solution==1, 1, 0)
rm(objective.in, const, rhs, direction, optimum)
```

Check expected statistics for new team
```{r, echo=FALSE}
sum(new_team[new_team$first_11_new == 1,]$now_cost)
sum(new_team$now_cost)
sum(new_team[new_team$first_11_new == 1,]$final_points) + 
  max(new_team[new_team$first_11_new == 1,]$final_points)
```

####################################################################################
#Chapter 7: Select ideal team
####################################################################################

Select best 11 for total amount (1000 + amount left over) minus 200
```{r, echo=FALSE}
#Strip nas 
players_2 <- players %>%
  filter(!is.na(final_points)) 
#Create objective function
objective.in <- players_2$final_points
#Create matrix for requirements
const <- t(as.matrix(tibble(value = players_2$now_cost,
                            goal_keeper = ifelse(players_2$position == "GK", 1, 0),
                            defender = ifelse(players_2$position == "DEF", 1, 0),
                            midfield = ifelse(players_2$position == "MID", 1, 0),
                            striker = ifelse(players_2$position == "FWD", 1, 0),
                            ones = rep(1, nrow(players_2)))))
#Create right hand side and inequality
rhs <- c(sum(previous_team$now_cost) - 200 + amount_left_over, 1, 3, 0, 1, 11)
direction  <- c("<=", "==", ">=", ">=", ">=", "==")
#Run, and make the team
optimum <-  lp(direction="max",  objective.in, const, direction,  rhs, all.bin = TRUE)
team_11 <- players_2[optimum$solution==1,] %>%
  mutate(first_11 = 1)
rm(objective.in, const, rhs, direction)
```

Select remaining 4 for the remaining balance
```{r, echo=FALSE}
#Remove selected players
players_remaining <- players_2[optimum$solution==0,]
#Create objective function
objective.in <- players_remaining$final_points
#Create matrix for requirements
const <- t(as.matrix(tibble(value = players_remaining$now_cost,
                            goal_keeper = ifelse(players_remaining$position == "GK", 1, 0),
                            defender = ifelse(players_remaining$position == "DEF", 1, 0),
                            midfield = ifelse(players_remaining$position == "MID", 1, 0),
                            striker = ifelse(players_remaining$position == "FWD", 1, 0),
                            ones = rep(1, nrow(players_remaining)))))
#Create right hand side and inequality. Don't go below 2 goalkeepers, 4 defenders and 2 strikers
rhs <- c(sum(previous_team$now_cost)+amount_left_over-sum(team_11$now_cost), 1, 1, 0, 0, 4)
direction  <- c("<=", "==", ">=", ">=", ">=", "==")
#Run, and make the team
optimum <-  lp(direction="max",  objective.in, const, direction,  rhs, all.bin = TRUE)
team_4 <- players_remaining[optimum$solution==1,] %>%
  mutate(first_11 = 0)
rm(objective.in, const, rhs, direction, players_remaining, optimum)
```

Bind together
```{r, echo=FALSE}
ideal_team <- rbind(team_11, team_4)
rm(team_11, team_4)
```

Check expected statistics
```{r, echo=FALSE}
sum(ideal_team[ideal_team$first_11 == 1,]$now_cost)
sum(ideal_team$now_cost)
sum(ideal_team[ideal_team$first_11 == 1,]$final_points) + 
  max(ideal_team[ideal_team$first_11 == 1,]$final_points)
```






